---
title: "Problem set 3"
output:
  html_document: default
  pdf_document: default
---

## Lily Shapiro, Csenge Petak

## Question 1:
We chose that species Staghorn coral (*Acropora cervicornis*), #215, because we wanted to investigate an endangered species and how one might go about saving one based on their population dynamics. First, we filtered the dataset for corals as coral reefs are known to be declining greatly due to global climate change. These stressors include thermal stress, sedimentation, pollution, and reduction in pH. Then, we looked at the papers pertaining to the matrices associated to corals in the COMADRE animal matrix database. We specifically chose the matrix for Staghorn corals since the article discussing this species investigates the demography of this threatened coral species and implications for its management and conservation, which exactly correlates with work discussed and conducted in this course. 

## Question 2:
We want to answer the following questions:

* What is the rate of population growth and what is the risk of extinction? 

* What are the factors that most contribute to population growth rate?

* What is the best way to restore the population if it would go to extinct otherwise?

## Question 3:

Eigenvalues:

```{r, warning = FALSE}
library(popbio)
load(file="comadre.RData")
matrix <- comadre$mat[215][[1]]$matA
e <- eigen(matrix)
a <- e$values
a[1]
```

Stable stage distribution:

```{r}
stable.stage(matrix)
```

Elasticity:

```{r}
elasticity(matrix)
```

Sensitivity:

```{r}
sensitivity(matrix)
```

## Question 4:

a. Dominant eigenvalue = 0.918, population is decreasing by about 8%. Computing the eigenvalue signifies a population decrease, which correlates with an increase in the number and size of threats affecting this particular coral species. Size of colonies and fragmentation (which may also be described as reproductive rate) was observed over 300 colonies for a period of two years. Because the eigenvalue is less than 1, the population will continually decrease, eventually possibly reaching extinction. The paper discusses viability analysis that predicts that populations will eventually reach quasi-extinction levels of 25% the original population size in less than 16 years (Mercado-Molina et al., 2015). 

b. Based on our elasticity and sensitivity analysis, we conclude that the survival rate of the large-sized coral colonies are most integral for population growth of this species. The stable-stage distribution calculation also showed that most of the population would consist of large-sized colonies (42%), rather than small or medium sizes. 

c. Conservation of large-sized coral colonies would be the most efficient way of maintaining the population. By halting the decline and increasing the rate of population growth focusing on maintaining the health and sustainability of large coral colonies, conservation of this species may be feasible. 


Citation: 
Mercado-Molina, A. E., Ruiz-Diaz, C. P., Pérez, M. E., Rodríguez-Barreras, R., & Sabat, A. M. (2015). Demography of the threatened coral Acropora cervicornis: implications for its management and conservation. Coral Reefs, 34(4), 1113–1124. https://doi.org/10.1007/s00338-015-1341-8


# Easton comments

- Really nice job on this assignment overall. You clearly communicate from the mathematical results (e.g. eigenvalue) and communicate what it means biologically. - It was smart including the .Rdata file in the same folder so when I downloaded the folder, I could also execute your .Rmd file
- Plots of the stable age distribution or of the elasticity values may have helped in interpretations.


